A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
design

>>>>

The R Project is a key tool for data analysis.  At the heart of the R
project is a \emph{vectorized, dynamic, lazy, functional, object-oriented}
programming language with a rather unusual combination of
features~\cite{ecoop12} designed to ease learning by non-programmer and
enable rapid development of new statistical methods.  The language was
designed~\cite{R96} as a successor to S~\cite{S88}.

In R, most data types are vectorized. Values are constructed by the
\c{c(...)} function: \c{c("hi","ho")} creates a vector of two strings. The
language does not differentiate scalars from vectors, thus \c{1==c(1)}. In
order to enable equational reasoning, R copies most values accessible
through multiple variables if they are written to.  Every linguistic
construct is desugared to a function call, even control flow statements,
assignments, and bracketing. Furthermore, all functions can be redefined in
packages or user code. This makes R both flexible and challenging to
compile~\cite{dls19}. In R, arguments to a user-defined function are bundled
into a thunk called a promise. Logically, a promise combines an expression's
code, its environment, and its value.  To access the value of a promise, one
must force it. Forcing a promise triggers evaluation and the computed value
is captured for future reference.  Values can be tagged by user-defined
attributes. For instance, one can attach the attribute \c{dim} to the value
\c{x<-c(1,2,3,4)} by evaluating \c{attr(x,"dim")<-c(2,2)}.  Once done,
arithmetic functions will treat \c x as a \c{2x2} matrix. Another attribute
is \c{class} which can be bound to a list of names. For instance,
\c{class(x)<-"human"}, sets the class of \c{x} to \c{human}.  Attributes are
used for object-oriented dispatch.

R supports reflection and meta-programming. The \c{substitute(e,envir)}
function yields the parse tree of the expression \c{e} after performing
substitutions defined by the bindings in \c{envir}.

\begin{lstlisting}
> substitute(expression(a + b), list(a = 1)))  
expression(1 + b)
\end{lstlisting}
\vspace{2mm}

\noindent
R allows programmatic manipulation of parse trees, which are themselves
first class objects. They are evaluated using the \c{eval(e,envir)}
function. Environment on the call stack can also be accessed.
